OpenMPI implementation of Burrows-Wheeler Transform for substring matching applied to genomics

### 2023-08-30 report/paper.tex
.

\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

@@

%% The amssymb package provides various useful mathematical symbols
\usepackage{hyperref}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{lipsum}

@@

@@

\begin{abstract}
%% Text of abstract
The aim of this project is to provide an efficient parallel implementation of a substring matching algorithm and apply it to the search of genes within the genome of an organism. After implementing the serial algorithm (using C/C++), an a priori study of the available parallelism has been conducted to identify the parts of the code that could be parallelized and then the code has been modified in order to run on multiple CPUs by using OpenMPI. The performance in terms of strong and weak scalability of the parallel implementation have been evaluated by running the code on Google Cloud Platform virtual instances in order to exploit the resources of a cluster of machines. A significant speedup has been obtained compared to the serial version, although under certain conditions the code is not able to scale efficiently as discussed later.
\end{abstract}

@@

\section{Biological context and Data}
\label{Biological context and Data}
The algorithm considered in this project is being used to find a gene within the genome of an organism. The genome is made up of DNA and contains the set of genes of an organism which are responsible for its traits and characteristics.
DNA is a long double helix molecule composed by a sequence of nucleotides\cite{nucleotide}, each nucleotide is composed by deoxyribose sugar, a phosphate group and a nucleobase. There are four types of nucleobases in DNA: guanine (G), adenine (A), cytosine (C) and thymine (T).
In bioinformatics applications, the genome and the genes are usually considered as long strings of computer code composed by four possible letters (G,A,C,T). \\
The genomes used for this analysis have been taken from the NCBI website\cite{ncbi}. \\
Genomes of different sizes have been used in order to study how the performance changes with respect to the input dataset dimension:
\begin{itemize}
    \item \emph{Escherichia Coli}, bacteria commonly found in the lower intestine of warm-blooded organisms\cite{ecoli}, 4.6MB \cite{ecoli dataset}
    \item \emph{Entamoeba Invadens}, an amoebozoa parasite of reptiles\cite{entamoeba}, 40.9MB \cite{entamoeba dataset}
    \item \emph{Formica Exsecta}, is a species of ant found from Western Europe to Asia \cite{formica}, 277.6MB \cite{formica dataset}
\end{itemize}

@@

\subsection{Burrows-Wheeler Transform}
The Burrows-Wheeler Transform restructures the original string in a way that is more compressible. It does so by building a matrix whose rows are all the cyclic shifts of the input string, then the rows are lexicographically sorted and finally the last column is selected as the output of the BWT.
Let's consider the following example in Figure \ref{fig:bwt-label} taken from \cite{mreasy}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/bwt.png}
    \caption{Burrows-Wheeler Transform}
    \label{fig:bwt-label}
\end{figure}
\\ We take the last column (BWT array) because it can be proven that we can recover all the cyclic rotations rows from it and it is the only column having this property that is important to compute the inverse of the transform \cite{geeks}. Moreover, it has better symbol clustering, which means that the same symbols are often grouped together and this favours efficient compression \cite{geeks}. \\
The implementation\cite{bwt.h me} of the Burrows-Wheeler Transform used in this project to compute the BWT array and the Suffix Array for a given genome string has been taken from \cite{geeks}.

\subsection{BWT pattern matching}
After computing the BWT array and the Suffix Array we can use them to efficiently find a pattern (gene) within the genome. The implementation provided in this project\cite{bwt.h me} and described here takes inspiration from \cite{mreasy} and was adapted to our scope using C++ language.\\
As it can be noticed in Figure \ref{fig:bwt-label}, in the first column of the previously computed matrix the nucleobases are sorted and grouped together, so it is possible to compute the bands (range of indexes of the Suffix Array) for all nucleobases. The algorithm starts from the last nucleobase in the pattern and keeps moving towards the start. At first the band of the last nucleobase is computed so we know where to start searching, then the idea is to reduce the band size at each iteration as we move backwards until we eventually either find the starting position of the pattern in the genome or we conclude that no match is found (if the band size reduces to zero before iterating all the pattern string). In order to do so we define a \emph{rank} function that tells how many times a character appeared in the BWT array before a certain index. Then, at each iteration over the nucleobases of the pattern, the start index and end index of the band are updated in the following way:
\begin{lstlisting}[language=c++]
char c = pattern[i];

@@

    \item \textbf{Suffix Array computation}: as it can be noticed by looking at the code implementation, it contains two for loops iterating over the genome string and an execution of the quicksort algorithm, so the computational cost of this phase can be computed as $$O(2*genomeSize + genomeSize*ln(genomeSize))$$

@@

@@

@@

@@

@@

\label{fig:times_table}

@@

@@

\section{A priori study of available parallelism}
The serial version of the code has been analyzed in order to understand which parts can be parallelized and what enhancement can be theoretically expected by executing the code on multiple CPUs cooperating with each other in parallel.\\
In order to estimate the improvement in performance, in terms of speedup, the \textbf{Amdalh's law} can be used:
$$ Speedup(N) = \frac{1}{S + \frac{P}{N}} $$
Where N is the number of nodes (virtual CPUs) that are exploited, S is the fraction of the code that cannot be parallelized and P is the fraction that can be parallelized. It can be noticed that obviously S+P=1 and the greater is P with respect to S the more speedup is obtained, at least theoretically.\\
The only parts that cannot be parallelized, at least not efficiently or simply, are the reading and preprocessing of the genome and gene files. Even the string matching algorithm based on the Burrows-Wheeler Transform is intrinsically serial, due to loop-carried dependencies, but it is possible to parallelize it by distributing chunks of the genome on different nodes so that the nodes can perform the BWT pattern search in parallel each on a different genome portion, smaller than the whole dataset. The choice of this strategy is also motivated by the fact that, as it can be noticed from the table in the previous section, the computational costs of the operations are highly dependent on the genome dataset size, so we can expect an enhancement by distributing it among different nodes. Whereas it is not possible to distribute the pattern string across different nodes because they all need to have the whole pattern to be able to find it. The methods and issues regarding the parallel implementation will be discussed in greater details in the next section. \\
The Amdalh's law has been applied after computing the S and P values for each dataset using the execution times in the table \ref{fig:times_table}. The expected theoretical speedups calculated for all datasets are shown in Figure \ref{fig:theoretical-speedup-label}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/theoretical_speedup.png}
    \caption{Theoretical Speedup}
    \label{fig:theoretical-speedup-label}
\end{figure}
\\It can be noticed that for larger datasets the expected speedup is lower than for smaller ones due to the non-parallelizable fraction S being greater (as shown in the execution times table \ref{fig:times_table}) probably because there are overheads due to the allocation and management of a bigger amount of memory.

@@

\subsubsection{Intra-regional}
The strong and weak scalability of the parallel implementation on a Fat Intra-regional cluster have been assessed using the speedup and efficiency formulas previously discussed and the results are shown in Figures \ref{fig:fat-intra-strong} and \ref{fig:fat-intra-weak}. \\
The speedup obtained in Figure \ref{fig:fat-intra-strong} can be considered a good result and it is in line with what was expected from the a priori analysis, it is only slightly lower due to the process synchronization and data transfer overheads that have a negative impact on the execution times especially for the case of larger datasets and many virtual cores. \\
Whereas, if we consider the weak scalability shown in Figure \ref{fig:fat-intra-weak}, we can notice that the efficiency drops faster for larger datasets, probably because the larger the dataset the more communication overhead is required to transfer it across the different nodes and also the memory allocation for the single node becomes more computationally expensive.

\subsubsection{Infra-regional}
The same experiments have been executed on a Fat Infra-regional cluster and as it can be noticed from Figures \ref{fig:fat-infra-strong} and \ref{fig:fat-infra-weak} the performance in terms of both strong and weak scalability is much worse than in the intra-regional case. \\
It is interesting to notice (Figure \ref{fig:fat-infra-strong}) that for smaller datasets, such as \emph{Escherichia Coli}, increasing the number of virtual cores beyond 8 vCores actually decreases the speedup; since 8 vCores is the capability of a single machine probably the reason for this strange behaviour is that just establishing the connection between machines far away from each other may require an overhead that is very large compared to both the data transfer overhead and the time needed to process the data in the case of a very small dataset. Even for medium and larger datasets we reach a maximum number of virtual cores after which the speedup decreases. Moreover, for all datasets the speedup obtained is lower than the intra-regional configuration given the same number of virtual cores. \\ 
This behaviour is confirmed by looking at the weak scalability plot (Figure \ref{fig:fat-infra-weak}) where we can notice that the efficiency drops much faster than the intra-regional case, because transfering data between machines located in different geographical regions demands a larger overhead. The efficiency drops faster for the smaller dataset for the reason discussed previously. \\
We can conclude that the intra-regional configuration is better in terms of both strong and weak scalability than the infra-regional cluster, that is not recommendable especially for small datasets.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/fat_intra_strong.png}
    \caption{Strong scalability for Fat Intra-regional cluster }
    \label{fig:fat-intra-strong}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/fat_intra_weak.png}
    \caption{Weak scalability for Fat Intra-regional cluster }
    \label{fig:fat-intra-weak}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/fat_infra_strong.png}
    \caption{Strong scalability for Fat Infra-regional cluster }
    \label{fig:fat-infra-strong}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/fat_infra_weak.png}
    \caption{Weak scalability for Fat Infra-regional cluster }
    \label{fig:fat-infra-weak}
\end{figure}

@@

\subsubsection{Intra-regional}
The results for strong and weak scalability on a Light Intra-region cluster are shown in Figures \ref{fig:light-intra-strong} and \ref{fig:light-intra-weak}. \\
As in the previous cases, strong scalability has been evaluated through the computation of the speedup (Figure \ref{fig:light-intra-strong}), which can be considered a good enhancement and is comparable to the results obtained for the fat intra-regional cluster. \\
Similarly to the fat intra-regional custer, also in this case it can be noticed that efficiency drops faster for larger datasets that are indeed less weakly scalable on intra-regional clusters compared to smaller ones (Figure \ref{fig:light-intra-weak}) since there is a greater overhead for transfering the data between different nodes.

\subsubsection{Infra-regional}
Finally, the same experiments have been conducted on the light infra-regional cluster. The results are shown in Figures \ref{fig:light-infra-strong} and \ref{fig:light-infra-weak}. We can observe a behaviour similar to the results obtained for the execution on the fat infra-regional cluster for both the strong and weak scalability. Indeed, the speedup is lower than the corresponding values for the light intra-regional cluster, especially for the smaller dataset, and it tends to decrease beyond a certain number of virtual cores. Also the efficiency (Figure \ref{fig:light-infra-weak}) results are similar to those obtained for the fat infra-regional cluster (Figure \ref{fig:fat-infra-weak}) showing that even in a light cluster the infra-regional configuration is less weakly scalable, especially for small datasets, compared to the intra-regional counterpart. 
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/light_intra_strong.png}
    \caption{Strong scalability for Light Intra-regional cluster }
    \label{fig:light-intra-strong}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/light_intra_weak.png}
    \caption{Weak scalability for Light Intra-regional cluster }
    \label{fig:light-intra-weak}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/light_infra_strong.png}
    \caption{Strong scalability for Light Infra-regional cluster }
    \label{fig:light-infra-strong}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/light_infra_weak.png}
    \caption{Weak scalability for Light Infra-regional cluster }
    \label{fig:light-infra-weak}
\end{figure}

>>>>

 \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
 \let\today\relax
\makeatletter
\def\ps@pprintTitle{%
    \let\@oddhead\@empty
    \let\@evenhead\@empty
    \def\@oddfoot{\footnotesize\itshape
         {} \hfill\today}%
    \let\@evenfoot\@oddfoot
    }
\makeatother

@@

%% The amssymb package provides various useful mathematical symbols
\usepackage{hyperref}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{lipsum}
\usepackage{float}

@@

\usepackage{caption}
\captionsetup{skip=2pt}

@@

\begin{abstract}
%% Text of abstract
The aim of this project is to provide an efficient parallel implementation of a substring matching algorithm and apply it to the search of genes within the genome of an organism. After implementing the serial algorithm (using C language), an a priori study of the available parallelism has been conducted to identify the parts of the code that could be parallelized and then the code has been modified in order to run on multiple CPUs by using OpenMPI. The performance in terms of strong and weak scalability of the parallel implementation have been evaluated by running the code on Google Cloud Platform virtual instances in order to exploit the resources of a cluster of machines. A significant speedup has been obtained compared to the serial version, although under certain conditions the code is not able to scale efficiently.
\end{abstract}

@@

\section{Biological context and Data}
\label{Biological context and Data}
The algorithm considered in this project is being used to find a gene within the genome of an organism. The genome is made up of DNA and contains the set of genes of an organism which are responsible for its traits and characteristics.
DNA is a long double helix molecule composed by a sequence of nucleotides, each nucleotide is composed by deoxyribose sugar, a phosphate group and a nucleobase\cite{nucleotide}. There are four types of nucleobases in DNA: guanine (G), adenine (A), cytosine (C) and thymine (T).
In bioinformatics applications, the genome and the genes are usually considered as long strings of computer code composed by four possible letters (G,A,C,T). \\
The genomes used for this analysis have been taken from the NCBI website\cite{ncbi}. \\
Genomes of different sizes have been used in order to study how the performance changes with respect to the input dataset dimension:
\begin{itemize}
    \item \emph{Escherichia Coli}, bacteria commonly found in the lower intestine of warm-blooded organisms\cite{ecoli}, 4.6MB \cite{ecoli dataset}
    \item \emph{Entamoeba Invadens}, an amoebozoa parasite of reptiles\cite{entamoeba}, 40.9MB \cite{entamoeba dataset}
    \item \emph{Formica Exsecta}, is a species of ant found from Western Europe to Asia \cite{formica}, 277.6MB \cite{formica dataset}
\end{itemize}

@@

\subsection{Burrows-Wheeler Transform}
The Burrows-Wheeler Transform restructures the original string in a way that is more compressible. It does so by building a matrix whose rows are all the cyclic shifts of the input string, then the rows are lexicographically sorted and finally the last column is selected as the output of the BWT.
Let's consider the following example in Figure \ref{fig:bwt-label} taken from the post on \emph{mr-easy.github.io}\cite{mreasy}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/bwt.png}
    \caption{Burrows-Wheeler Transform}
    \label{fig:bwt-label}
\end{figure}
\\ We take the last column (BWT array) because it can be proven that we can recover all the cyclic rotations rows from it and it is the only column having this property that is important to compute the inverse of the transform \cite{geeks}. Moreover, it has better symbol clustering, which means that the same symbols are often grouped together and this favours efficient compression \cite{geeks}. \\
The implementation\cite{bwt.h me} of the Burrows-Wheeler Transform used in this project to compute the BWT array and the Suffix Array for a given genome string has been taken from the article on \emph{GeeksforGeeks}\cite{geeks}.

\subsection{BWT pattern matching}
After computing the BWT array and the Suffix Array we can use them to efficiently find a pattern (gene) within the genome. The implementation provided in this project\cite{bwt.h me} and described here takes inspiration from \emph{mr-easy.github.io}'s article\cite{mreasy} and was adapted to our scope using C language.\\
As it can be noticed in Figure \ref{fig:bwt-label}, in the first column of the previously computed matrix the nucleobases are sorted and grouped together, so it is possible to compute the bands (range of indexes of the Suffix Array) for all nucleobases. The algorithm starts from the last nucleobase in the pattern and keeps moving towards the start. At first the band of the last nucleobase is computed so we know where to start searching, then the idea is to reduce the band size at each iteration as we move backwards until we eventually either find the starting position of the pattern in the genome or we conclude that no match is found (if the band size reduces to zero before iterating all the pattern string). In order to do so we define a \emph{rank} function that tells how many times a character appeared in the BWT array before a certain index. Then, at each iteration over the nucleobases of the pattern, the start index and end index of the band are updated in the following way:
\begin{lstlisting}[language=c++]
char c = pattern[i];

@@

    \item \textbf{Suffix Array computation}: as it can be noticed by looking at the code implementation, it contains two separate for loops iterating over the genome string and an execution of the quicksort algorithm, so the computational cost of this phase can be computed as $$O(2*genomeSize + genomeSize*ln(genomeSize))$$

@@

\begin{table}[H]

@@

\caption{Computational costs for each step of the code}

@@

\end{table}

@@

\caption{Genome and pattern sizes for the datasets used}
\vspace{-2mm}%Put here to reduce too much white space after your table 
\end{table}

@@

@@

\caption{Genome and pattern sizes for the datasets used}
\vspace{-2mm}%Put here to reduce too much white space after your table 
\label{table:times_table}
\end{table}

@@

\section{A priori study of available parallelism}
The serial version of the code has been analyzed in order to understand which parts can be parallelized and what enhancement can be theoretically expected by executing the code on multiple CPUs cooperating with each other in parallel.\\
In order to estimate the improvement in performance, in terms of speedup, the \textbf{Amdalh's law} can be used:
$$ Speedup(N) = \frac{1}{S + \frac{P}{N}} $$
Where N is the number of nodes (virtual CPUs) that are exploited, S is the fraction of the code that cannot be parallelized and P is the fraction that can be parallelized. It can be noticed that obviously S+P=1 and the greater is P with respect to S the more speedup is obtained, at least theoretically.\\
The only parts that cannot be parallelized, at least not efficiently or simply, are the reading and preprocessing of the genome and gene files. Even the string matching algorithm based on the Burrows-Wheeler Transform is intrinsically serial, due to loop-carried dependencies, but it is possible to parallelize it by distributing chunks of the genome on different nodes so that the nodes can perform the BWT pattern search in parallel each on a different genome portion, smaller than the whole dataset. The choice of this strategy is also motivated by the fact that, as it can be noticed from the table in the previous section, the computational costs of the operations are highly dependent on the genome dataset size, so we can expect an enhancement by distributing it among different nodes. Whereas it is not possible to distribute the pattern string across different nodes because they all need to have the whole pattern to be able to find it. The methods and issues regarding the parallel implementation will be discussed in greater details in the next section. \\
The Amdalh's law has been applied after computing the S and P values for each dataset using the execution times in the table \ref{table:times_table}. The expected theoretical speedups calculated for all datasets are shown in Figure \ref{fig:theoretical-speedup-label}.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{images/theoretical_speedup.png}
    \caption{Theoretical Speedup}
    \label{fig:theoretical-speedup-label}
\end{figure}
It can be noticed that for larger datasets the expected speedup is lower than for smaller ones due to the non-parallelizable fraction S being greater (as shown in the execution times table \ref{table:times_table}) probably because there are overheads due to the allocation and management of a bigger amount of memory.

@@

\subsubsection{Intra-regional}
The strong and weak scalability of the parallel implementation on a Fat Intra-regional cluster have been assessed using the speedup and efficiency formulas previously discussed and the results are shown in Figures \ref{fig:fat-intra-strong} and \ref{fig:fat-intra-weak}. \\
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/fat_intra_strong.png}
    \caption{Strong scalability for Fat Intra-regional cluster }
    \label{fig:fat-intra-strong}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/fat_intra_weak.png}
    \caption{Weak scalability for Fat Intra-regional cluster }
    \label{fig:fat-intra-weak}
\end{figure}
The speedup obtained in Figure \ref{fig:fat-intra-strong} can be considered a good result and it is in line with what was expected from the a priori analysis, it is only slightly lower due to the process synchronization and data transfer overheads that have a negative impact on the execution times especially for the case of larger datasets and many virtual cores. \\
Whereas, if we consider the weak scalability shown in Figure \ref{fig:fat-intra-weak}, we can notice that the efficiency drops faster for larger datasets, probably because the larger the dataset the more communication overhead is required to transfer it across the different nodes and also the memory allocation for the single node becomes more computationally expensive.

\subsubsection{Infra-regional}
The same experiments have been executed on a Fat Infra-regional cluster and as it can be noticed from Figures \ref{fig:fat-infra-strong} and \ref{fig:fat-infra-weak} the performance in terms of both strong and weak scalability is much worse than in the intra-regional case. \\
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/fat_infra_strong.png}
    \caption{Strong scalability for Fat Infra-regional cluster }
    \label{fig:fat-infra-strong}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/fat_infra_weak.png}
    \caption{Weak scalability for Fat Infra-regional cluster }
    \label{fig:fat-infra-weak}
\end{figure}
It is interesting to notice (Figure \ref{fig:fat-infra-strong}) that for smaller datasets, such as \emph{Escherichia Coli}, increasing the number of virtual cores beyond 8 vCores actually decreases the speedup; since 8 vCores is the capability of a single machine probably the reason for this strange behaviour is that just establishing the connection between machines far away from each other may require an overhead that is very large compared to both the data transfer overhead and the time needed to process the data in the case of a very small dataset. Even for medium and larger datasets we reach a maximum number of virtual cores after which the speedup decreases. Moreover, for all datasets the speedup obtained is lower than the intra-regional configuration given the same number of virtual cores. \\ 
This behaviour is confirmed by looking at the weak scalability plot (Figure \ref{fig:fat-infra-weak}) where we can notice that the efficiency drops much faster than the intra-regional case, because transfering data between machines located in different geographical regions demands a larger overhead. The efficiency drops faster for the smaller dataset for the reason discussed previously. \\
We can conclude that the intra-regional configuration is better in terms of both strong and weak scalability than the infra-regional cluster, that is not recommendable especially for small datasets.

@@

\subsubsection{Intra-regional}
The results for strong and weak scalability on a Light Intra-region cluster are shown in Figures \ref{fig:light-intra-strong} and \ref{fig:light-intra-weak}. \\
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/light_intra_strong.png}
    \caption{Strong scalability for Light Intra-regional cluster }
    \label{fig:light-intra-strong}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/light_intra_weak.png}
    \caption{Weak scalability for Light Intra-regional cluster }
    \label{fig:light-intra-weak}
\end{figure}
As in the previous cases, strong scalability has been evaluated through the computation of the speedup (Figure \ref{fig:light-intra-strong}), which can be considered a good enhancement and is comparable to the results obtained for the fat intra-regional cluster. \\
Similarly to the fat intra-regional custer, also in this case it can be noticed that efficiency drops faster for larger datasets that are indeed less weakly scalable on intra-regional clusters compared to smaller ones (Figure \ref{fig:light-intra-weak}) since there is a greater overhead for transfering the data between different nodes.

\subsubsection{Infra-regional}
Finally, the same experiments have been conducted on the light infra-regional cluster. The results are shown in Figures \ref{fig:light-infra-strong} and \ref{fig:light-infra-weak}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/light_infra_strong.png}
    \caption{Strong scalability for Light Infra-regional cluster }
    \label{fig:light-infra-strong}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{images/light_infra_weak.png}
    \caption{Weak scalability for Light Infra-regional cluster }
    \label{fig:light-infra-weak}
\end{figure}
We can observe a behaviour similar to the results obtained for the execution on the fat infra-regional cluster for both the strong and weak scalability. Indeed, the speedup is lower than the corresponding values for the light intra-regional cluster, especially for the smaller dataset, and it tends to decrease beyond a certain number of virtual cores. Also the efficiency (Figure \ref{fig:light-infra-weak}) results are similar to those obtained for the fat infra-regional cluster (Figure \ref{fig:fat-infra-weak}) showing that even in a light cluster the infra-regional configuration is less weakly scalable, especially for small datasets, compared to the intra-regional counterpart.
